---
title: "airstats"
output: html_document
editor_options: 
  chunk_output_type: console
chunk_output_type: message=FALSE, warning=FALSE, include=FALSE
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Loading the libraries
```{r}
library(usmap)
library(sf) #not sure which mapping to use for future use
library(tidyverse)
library(bbplot)
```

# Importing the data
```{r}
airports <- read.csv("airports.csv")
countries <- read.csv("countries.csv")
regions <- read.csv("regions.csv")
```

# Having a quick look
```{r}
glimpse(airports)
```

# Grabbing data from other tables, selecting the columns I want, and renaming a few
```{r}
data <- left_join(airports, countries, by = c("iso_country" = "code")) %>% 
  select(ident, 
         type, 
         name.x, 
         latitude_deg, 
         longitude_deg, 
         elevation_ft, 
         iso_country, 
         scheduled_service, 
         name.y,
         iso_region) %>% 
  rename(airport_name = name.x,
         latitude = latitude_deg,
         longitude = longitude_deg,
         country = name.y)
```

# Checking I have everything I need
```{r}
glimpse(data)
```

# Tidying up airport type names
```{r}
data$type[data$type=="balloonport"] <- "Balloonport"
data$type[data$type=="large_airport"] <- "Large Airport"
data$type[data$type=="medium_airport"] <- "Medium Airport"
data$type[data$type=="seaplane_base"] <- "Seaplane Base"
data$type[data$type=="closed"] <- "Closed Airport"
data$type[data$type=="heliport"] <- "Heliport"
data$type[data$type=="small_airport"] <- "Small Airport"
```

# Adding elevations into 1000ft groupings
```{r}
labs <- c("0-999","1000-1999","2000-2999","3000-3999","4000-4999","5000-5999","6000+")

data$elevation_group <- cut(data$elevation_ft, breaks = c(seq(0,6000, by = 1000), Inf), labels = labs, right = FALSE)
```

# Looking at airport types in USA
```{r}
data %>% 
  filter(country %in% c("United States")) %>% 
  #ggplot(aes(x = fct_rev(fct_infreq(type)))) + #fct_rev changes order
  ggplot(aes(x = fct_infreq(type))) + #fct_rev changes order
  geom_bar(stat = "count", fill="#1380A1") +
  ylim(c(0,15000)) +
  geom_text(stat='count', aes(label=..count..),
             hjust = 0.5, 
             vjust = -0.3, 
             colour = "black", 
             fill = NA, 
             label.size = NA, 
             family="Helvetica", 
             size = 5) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  #coord_flip() +
  #scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  bbc_style() +
  labs(title="U.S Airport Types",
       subtitle = "Different types of airports in the U.S.A",
       x = "Airport Type",
       y = "Count")
```

# Mapping airport types in Alaska, USA
```{r}
data %>% 
  filter(iso_region %in% c("US-AK")) %>%
  count(type, longitude, latitude, sort = TRUE) %>% 
  ggplot(aes(longitude, latitude, colour = type)) +
  geom_point() +
  lims(x = c(-180, -130), y = c(51, 72)) +
  scale_size_continuous() +
  labs(title="Airport Types in Alaska, US") +
  borders(
  database = "world",
  regions = ".",
  fill = NA,
  colour = "grey50",
  xlim = NULL,
  ylim = NULL) +
  bbc_style() +
  theme(legend.position = "right")
```

# Investigating heliports seen in the North Sea
```{r}
data %>% 
  #filter(type %in% c("Heliport")) %>%
  count(type, longitude, latitude, sort = TRUE) %>% 
  ggplot(aes(longitude, latitude, colour = type)) +
  geom_point() +
  lims(x = c(-12, 15), y = c(48, 63)) +
  scale_size_continuous() +
  labs(title="Heliports in the North Sea",
       subtitle="Heliports used as a commuter service to oil rigs") +
  borders(
  database = "world",
  regions = ".",
  fill = NA,
  colour = "grey50",
  xlim = NULL,
  ylim = NULL) +
  bbc_style() +
  theme(legend.position = "right")
```

# Charting splits of elevation across airport types
```{r}
data %>% 
  drop_na(elevation_group) %>% 
  ggplot(aes(x = fct_infreq(type), fill=forcats::fct_rev(elevation_group))) +
  geom_bar(stat = "count", position = "fill") +
  scale_fill_manual(values = c("#CCE9EF","#99D4DF","#66BECF","#33A9BF","#1A8FA6","#005869","#003B46")) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  bbc_style() +
  theme(legend.position = "right") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  labs(title="Elevation of different airport types")
```

# Looking at iso_region for US 
```{r}
data %>% 
  filter(country %in% c("United States")) %>% 
  count(iso_region, sort = TRUE)
```

```{r}
plot_usmap(regions = "states") + 
  labs(title = "U.S. States",
       subtitle = "This is a blank map of the United States.") + 
  theme(panel.background=element_blank())
```
